A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-09-27 paper/paper.tex
paper: minor polish

% Simplicial and Cubical sets
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{\square}}}
\renewcommand{\I}[1]{\ensuremath{\mathsf{I}^{#1}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

@@

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\newcommandx{\restrf}[1][1]{\restr{frame}[][#1][][][][][]}

@@

\begin{document}
\thispagestyle{empty}
\maketitle
\begin{abstract}
  We present a parametricity-based construction of semi-simplicial sets and semi-cubical sets in indexed form, in the calculus of inductive constructions. The accompanying artifact is an axiom-less~\footnote{Technically speaking, we use some axioms related to functional extensionality, but the requirement arises only from our generalization to $\nu$-types; if $\nu$ is finite (say, unit type corresponding to augmented semi-simplicial types), this requirement vanishes.} Coq formalization that uses sophisticated rewriting techniques and pushes the boundaries of proof assistant technology.
\end{abstract}
\tableofcontents
\newpage

\section{Overture}

In the context of typed functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch14,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of (Kreisel) realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation from Hugo Moeneclaey}.

@@

@@

\subsection{Generalization to \texorpdfstring{$\nu$}{nu}}
\begin{table}[H]
  \begin{tabularx}{\linewidth}{p{.2\linewidth}|p{.4\linewidth}|p{.3\linewidth}}
    \toprule
    Value of $\nu$ & 1                     & 2                  \\
    \graymidrule
    Interpretation & Semi-simplicial types & Semi-cubical types \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

\begin{align*}
  \mathsf{Dom} & : \mathsf{Type}                                          \\
  \mathsf{UIP} & : \forall x y : \mathsf{Dom}, \forall p q : x = y, p = q
\end{align*}

>>>>

% Simplicial and Cubical sets
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{\square}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

@@

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\newcommandx{\restrf}[1][1]{\restr{frame}[][#1][][][][][]}

@@

\begin{document}
\maketitle
\begin{abstract}
  We present a parametricity-based construction of semi-simplicial sets and semi-cubical sets in indexed form, in the calculus of inductive constructions. The accompanying artifact is an axiom-less~\footnote{Technically speaking, we use some axioms related to functional extensionality, but the requirement arises only from our generalization to $\nu$-types; if $\nu$ is finite (say, unit type corresponding to augmented semi-simplicial types), this requirement vanishes.} Coq formalization that uses sophisticated rewriting techniques and pushes the boundaries of proof assistant technology.
\end{abstract}
\tableofcontents
\newpage

\section{Overture}
In the context of typed functional programming, Reynolds' parametricity~\cite{reynolds72} interprets types as relations characterizing the observational equality of programs of this type. Parametricity can be iterated, and it has been noted that iterated Reynolds' parametricity has a cubical flavor~\cite{johann17,altenkirch14,moulin16}. We obtain a \emph{unary} variant of Reynolds' \emph{binary} parametricity by using predicates instead of relations, in which case we obtain a form of (Kreisel) realizability~\cite{bernardy12,moulin16}. It has then been noted that iterated unary parametricity has an augmented simplicial flavor~\footnote{Observation from Hugo Moeneclaey}.

@@

\begin{table}[H]
  \begin{tabularx}{\linewidth}{p{.2\linewidth}|p{.4\linewidth}|p{.3\linewidth}}
    \toprule
    Value of $\nu$ & 1                               & 2                  \\
    \graymidrule
    Interpretation & Augmented semi-simplicial types & Semi-cubical types \\
    \bottomrule
  \end{tabularx}
\end{table}

@@

@@

\begin{align*}
  \textsf{Dom} & : \textsf{Type}                                          \\
  \textsf{UIP} & : \forall x y : \textsf{Dom}, \forall p q : x = y, p = q
\end{align*}
